Bayesian calibration of a hybrid nuclear collision model\\ using $\boldsymbol p$-Pb and Pb-Pb data from the LHC

### 2019-05-07 paper/response.txt

Thank you for the constructive criticisms and comments. We have significantly revised the manuscript to improve its clarity and address several of your concerns.

Response to the specific concerns of the referee:

"A parton model of the proton has direct consequences not only for heavy-ion collisions, but also for p+p collisions. The authors should present predictions for p+p scattering using the most probable parameters (or using the posterior distribution of parameters w, v, n_c) for the differential elastic p+p scattering cross section. The prediction can be compared to parametrized data at 5 TeV. As a proxy for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019 can be used."

@@

Thank you for pointing this out. We've added the following comment to explain the truncation of the constituent number prior:

One exception is the constituent number $n_c$ which we limit for practical considerations. Recall that each constituent fluctuates independently, weighted by a gamma random variable. Hence for constituent numbers $n_c > 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish. To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases. Eventually, these required fluctuations become unreasonably large. We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.

"For the model predictions from the posterior experimental data (even preliminary) should be added, whenever available."

We've been careful to include all available data (public or otherwise), but it is possible that we missed a dataset. In response to your suggestion, we sent an email to the ALICE collaboration to request their preliminary data, e.g. Pb-Pb mean pT, which could be added to the posterior predictions where data is missing. Unfortunately, it is not their policy to share preliminary data, and our request was denied. If you can direct us to a specific dataset, we would be happy to include it in the maximum a posteriori observable plot.

### 2019-05-07 paper/paper.tex
Final revisions

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\sqrts}{\sqrt{s_\mathrm{NN}}}
\newcommand{\fmc}{\ensuremath{\text{fm}/c}}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\sigmaf}{\sigma_\text{fluct}}
\newcommand{\taufs}{\tau_\text{fs}}
\newcommand{\dmin}{d_\text{min}}
\newcommand{\Tsw}{T_\text{switch}}
\newcommand{\T}{\tilde{T}}

\newcommand{\smin}{{(\eta/s)_\mathrm{min}}}
\newcommand{\sslope}{{a_\mathrm{shear}}}
\newcommand{\scrv}{{c_\mathrm{shear}}}
\newcommand{\bmax}{{(\zeta/s)_\mathrm{max}}}
\newcommand{\bloc}{{T_\text{bulk}}}
\newcommand{\bwidth}{{\Delta T_\text{bulk}}}

@@

Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-femtometer length scales becomes important to describe bulk properties of the produced system.

One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a given theory framework can be validated by testing its predictions for multiple collision systems, using a single self-consistent set of model parameters where only the nuclear structure in the model is permitted to vary.

@@

Several such observables have been identified in proton-proton and proton-lepton scattering data.
Measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV, for instance, found an unexpected dip in the inelasticity density of $p$-$p$ collisions at zero impact parameter \cite{Antchev:2011zz}.
It was later realized that this depression, or so-called hollowness effect in the $p$-$p$ inelastic collision profile \cite{Arriola2016}, can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.

@@

With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin in Sec.~\ref{sec:model} by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics, and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged-particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
The sheets of colliding nuclear density penetrate and pass through each other in time $\tau_\text{overlap} \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
The resulting nuclear overlap time $\tau_\text{overlap} \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is reasonable to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying density inside each nucleus.

@@

In this work, we pursue a less restrictive and more parametric description of the nucleon where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
\begin{equation}
  \label{eq:rho}
  \rho_{A, B}(\xv) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\xv - \xv_i),
\end{equation}
where each constituent density $\rho_c$ is described by a Gaussian distribution
\begin{equation}
  \rho_c(\xv) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 v^2}\right)
\end{equation}
of variable width $v$.
The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution
\begin{equation}
  \label{eq:radial_dist}
  \rho_\text{pos}(\xv_i; \xv_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_{A,B}|^2}{2 r^2}\right),
\end{equation}
where $r$ is a parameter that varies the sampling radius of the constituent positions $\xv_i$ about the nucleon position $\xv_{A,B}$.
As a matter of convenience, we sample the nucleon positions \emph{before} determining the constituent positions.
This leads to small discrepancies between the designated nucleon positions $\xv_{A,B}$ and the actual position of each nucleon's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
The parameter $r$ should thus be interpreted with care.
It is a computational sampling radius, \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.

@@

Hereafter, we switch to Milne coordinates $(\xv_\perp, \eta_s, \tau)$, where $\xv_\perp$ is a Cartesian coordinate $(x, y)$ in the transverse plane, $\eta_s = \tfrac{1}{2} \log[(t+z)/(t-z)]$ is the system's spacetime rapidity, and $\tau = \sqrt{t^2 - z^2}$ its proper time.
Assuming the nucleons collide at the sampled impact parameter $\bv$, we assign each nucleon a participant thickness
\begin{equation}
  \label{eq:fluctuated_thick}
  \T_{A, B}(\xv_\perp) \equiv \int dz\, \tilde{\rho}_{A,B}(\xv_\perp \pm \bv/2, z),
\end{equation}
which projects its \emph{fluctuated} nucleon density
\begin{equation}
  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c(\xv - \xv_i)
\end{equation}
onto the transverse plane $\xv_\perp$ centered at midrapidity.
This fluctuated nucleon density $\tilde{\rho}_{A,B}$ equals the nucleon density $\rho_{A,B}$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
Such \emph{ad hoc} random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.

The resulting nucleon fluctuation variance naturally falls like ${\sim}1/ n_c$, where $n_c$ is the number of independent fluctuating sources.
This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
We therefore reparametrize the constituent fluctuations using
\begin{equation}
  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
\end{equation}
where $k$ is the shape parameter (inverse variance) of the gamma distribution weights and $n_c$ is the number of nucleon constituents.

@@

A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times M_p(\T_A, \T_B),
\end{equation}
where $M_p$ is a family of functions known as the generalized means
\begin{equation}
  M_p(x, y) = \left( \frac{x^p + y^p}{2} \right)^{1/p}.
\end{equation}
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  M_p(x, y) =
  \begin{cases}
    \max(x, y) & p \rightarrow +\infty, \\[\extraspace]
    (x + y)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{x y} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, x y/(x + y) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(x, y) & p \rightarrow -\infty.
  \end{cases}
  \label{eq:trento_p}
\end{equation}

Note, the form of Eq.~\eqref{eq:gmean} differs somewhat from our previous studies \cite{Moreland:2014oya, Bernhard:2016tnd}, which parametrized the system's \emph{entropy} density using the generalized mean ansatz:
\begin{equation}
  s(\xv_\perp, \eta_s=0, \tau_0) \propto M_p(\T_A, \T_B).
\end{equation}
Our motivation for reinterpretting the left-side of this equation as an energy density in the present work is to enable the application of pre-equilibrium free streaming equations (described shortly) which require the system's initial energy density as input.
This modification is permissible since the \trento\ model is based on particle yield observations coupled with the approximate scaling relation $d\nch/d\eta \propto M_p(\T_A, \T_B)$, which remains valid for both static entropy density and free streamed energy density initialization.

\begin{figure}
  \begin{tikzpicture}
    \node {\includegraphics{thickness}};
    \node[label=above:{10 fm}] (a) at (0, 1.1) {};
    \node (b) at (0, -1.1) {};
    \draw [<->, semithick] (a) to (b);
  \end{tikzpicture}
  \caption{\label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus. The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
\end{figure}

Up to this point we have restricted our attention to a single nucleon-nucleon collision.
Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
The \trento\ model readily generalizes from individual nucleon-nucleon collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
Figure~\ref{fig:thickness} shows the effect of adding nucleon substructure to a generic lead nucleus.
Additional fluctuations emerge over the length scale of a nucleon, but the macroscopic geometry of the nucleus is largely unchanged.

@@

\begin{figure}
  \includegraphics{coupling}
  \caption{Cartoon of the free streaming approximation for hydrodynamic initialization. The initial state is free streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling). This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
  \label{fig:coupling}
\end{figure}

@@

The parametric energy deposition ansatz in Eq.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
\begin{equation}
  \label{freestream}
  p^\mu \partial_\mu f(x, p) = 0,
\end{equation}
through its dependence on the underlying distribution function $f(x, p)$.
Equation~\eqref{freestream}, however, simplifies for a boost-invariant gas of massless noninteracting partons with locally isotropic transverse momentum distributions.
Subject to these assumptions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ equals \cite{Broniowski:2008qk, Liu:2015nwa}
\begin{multline}
  \label{energy-momentum}
  T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau) =\\
  \frac{\tau_0}{\tau} \int_{-\pi}^{\pi} d\phi_p\, \hat{p}^\mu \hat{p}^\nu e(\xv_\perp - (\tau - \tau_0)\hat{\mathbf{p}}_\perp, \eta_s=0, \tau_0),
\end{multline}
where $\hat{p}^\mu = (1, \cos \phi_q, \sin \phi_q, 0)$ and $\hat{\mathbf{p}}_\perp = (\cos \phi_q, \sin \phi_q)$ are momentum unit vectors.
Here we assume that the time $\tau_0 \ll 1~\fmc$ is small and define $\taufs \equiv \tau - \tau_0 \approx \tau$.
We also combine the unknown constant in Eq.~\eqref{eq:gmean} with the pre-factor $\tau_0$ in Eq.~\eqref{energy-momentum} to produce a single normalization factor $\text{Norm} = \text{const} \times \tau_0$ with units of energy.
The solution \eqref{energy-momentum} is then decomposed in hydrodynamic form
\begin{equation}
  \label{hydro-eqn}
  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu},
\end{equation}
where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} = g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue
\begin{equation}
  T^{\mu\nu} u_\nu = e u^\mu.
\end{equation}
The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eq.~\eqref{hydro-eqn}
\begin{align}
  \Pi &= -\frac{1}{3} \mathrm{Tr}(\Delta_{\mu\nu} T^{\mu\nu}) - P,\\
  \pi^{\mu\nu} &= T^{\mu\nu} - e u^\mu u^\nu + (P + \Pi) \Delta^{\mu\nu}.
\end{align}

@@

\begin{figure}[t]
  \includegraphics{viscosity_dof}
  \caption{Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations. Lines are chosen for illustrative purposes only and do not represent all possible variability. For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.}
  \label{fig:viscosity_dof}
\end{figure}

@@

The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\ym$ describe experimental measurements $\ye$.
The problem involves three distinct components:
\begin{enumerate}[itemsep=0pt, leftmargin=2\parindent]
  \item $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
  \item $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
  \item $E$: the evidence provided by the experimental data $\ye$ and its corresponding uncertainties.
\end{enumerate}
As a practical matter, we always assume that hypothesis $H_f$ is correct, meaning there are no glaring flaws in our chosen theoretical framework.
This is a significant assumption; the application of hydrodynamic simulations to small collision systems is speculative, and our conclusions are conditional on the framework making sense.

>>>>

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\sqrts}{\sqrt{s_\mathrm{NN}}}
\newcommand{\fmc}{\ensuremath{\text{fm}/c}}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\sigmaf}{\sigma_\text{fluct}}
\newcommand{\taufs}{\tau_\text{fs}}
\newcommand{\dmin}{d_\text{min}}
\newcommand{\Tsw}{T_\text{switch}}
\newcommand{\Tc}{T_c}
\newcommand{\T}{\tilde{T}}

\newcommand{\rc}{r_{cp}}
\newcommand{\nc}{n_c}
\newcommand{\wc}{w_c}

\newcommand{\smin}{{(\eta/s)_\mathrm{min}}}
\newcommand{\sslope}{{(\eta/s)_\mathrm{slope}}}
\newcommand{\scrv}{{(\eta/s)_\mathrm{crv}}}
\newcommand{\bmax}{{(\zeta/s)_\mathrm{max}}}
\newcommand{\bwidth}{{(\zeta/s)_\mathrm{width}}}
\newcommand{\bloc}{{(\zeta/s)_{T_\mathrm{peak}}}}

@@

Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-femtometer length scales becomes important to describe bulk properties of the produced system.

One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a proposed theory framework can be validated by testing its predictions for multiple collision systems using a single self-consistent set of model parameters, where only the nuclear structure in the model is permitted to vary.

@@

Several such observables have been identified in proton-proton and proton-lepton scattering data.
Measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV, for instance, found an unexpected dip in the inelasticity density of $p$-$p$ collisions at zero impact parameter \cite{Antchev:2011zz}.
It was later realized that this depression, or so-called hollowness effect, in the $p$-$p$ inelastic collision profile \cite{Arriola2016} can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.

@@

With these considerations in mind, we present progress toward a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
We begin in Sec.~\ref{sec:model} by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics, and microscopic Boltzmann transport.
In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged-particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.

@@

The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
The sheets of colliding nuclear density penetrate and pass through each other in time $\tau_\text{overlap} \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
The resulting nuclear overlap time $\tau_\text{overlap} \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is reasonable to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying transverse density inside each nucleus.

@@

In this work, we pursue a less restrictive and more parametric description of the nucleon where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $\nc$ independent constituents
\begin{equation}
  \label{eq:rho}
  \rho_{A, B}(\xv) = \frac{1}{\nc} \sum\limits_{i=1}^{\nc} \rho_c(\xv - \xv_i),
\end{equation}
where each constituent density $\rho_c$ is described by a Gaussian distribution
\begin{equation}
  \rho_c(\xv) = \frac{1}{(2 \pi \wc^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 \wc^2}\right)
\end{equation}
of variable width $\wc$.
The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian probability distribution
\begin{equation}
  \label{eq:radial_dist}
  P(\xv_i) = \frac{1}{(2\pi \rc^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_n|^2}{2\, \rc^2}\right),
\end{equation}
where $\rc$ is a free parameter that varies the dispersion of the constituent positions $\xv_i$ about each nucleon position $\xv_n$.
As a matter of convenience, we sample the nucleon positions \emph{before} determining the constituent positions.
This leads to small discrepancies between the designated nucleon positions and the actual position of each nucleon's center-of-mass, owing to fluctuations in the constituent positions.
The parameter $\rc$ should thus be interpreted with care.
It is a computational sampling radius, \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.

@@

Hereafter, we switch to Milne coordinates $(\xv_\perp, \eta_s, \tau)$, where $\xv_\perp$ is a Cartesian coordinate $(x, y)$ in the transverse plane, $\eta_s = \tfrac{1}{2} \log[(t+z)/(t-z)]$ is the system's spacetime rapidity, and $\tau = \sqrt{t^2 - z^2}$ its proper time.
Assuming the nucleons collide at the sampled impact parameter $\bv$, we assign each nucleon a participant thickness
\begin{equation}
  \label{eq:fluctuated_thick}
  \T_{A, B}(\xv_\perp) \equiv \int dz\, \tilde{\rho}_{A,B}(\xv_\perp \pm \bv/2, z),
\end{equation}
which projects its \emph{fluctuated} nucleon density
\begin{equation}
  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{\nc} \sum\limits_{i=1}^{\nc} \gamma_i\, \rho_c(\xv - \xv_i)
\end{equation}
onto the transverse plane $\xv_\perp$ centered at midrapidity.
This fluctuated nucleon density $\tilde{\rho}_{A,B}$ equals the nucleon density $\rho_{A,B}$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
Such \emph{ad hoc} random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.

The resulting nucleon fluctuation variance naturally falls like ${\sim}1/ \nc$, where $\nc$ is the number of constituents.
This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger.
We therefore reparametrize the constituent fluctuations using
\begin{equation}
  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, \nc},
\end{equation}
where $k$ is the shape parameter (inverse variance) of the gamma distribution weights and $\nc$ is the number of nucleon constituents.

@@

A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times M_p(\T_A, \T_B),
\end{equation}
where $M_p$ is a family of functions known as the generalized means
\begin{equation}
  M_p(x, y) = \left( \frac{x^p + y^p}{2} \right)^{1/p}.
\end{equation}
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
For certain discrete $p$ values, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  M_p(x, y) =
  \begin{cases}
    \max(x, y) & p \rightarrow +\infty, \\[\extraspace]
    (x + y)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{x y} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, x y/(x + y) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(x, y) & p \rightarrow -\infty.
  \end{cases}
  \label{eq:trento_p}
\end{equation}

Note, the form of Eq.~\eqref{eq:gmean} differs somewhat from our previous studies \cite{Moreland:2014oya, Bernhard:2016tnd}, which parametrized the system's \emph{entropy} density using the generalized mean ansatz:
\begin{equation}
  s(\xv_\perp, \eta_s=0, \tau_0) \propto M_p(\T_A, \T_B).
\end{equation}
Our motivation for reinterpreting the left-side of this equation as an energy density in the present work is to enable the application of pre-equilibrium free streaming equations (described shortly) which require the system's initial energy density as input.
This modification is permissible since the \trento\ model is formulated using particle yield observations and the approximate scaling relation $d\nch/d\eta \propto M_p(\T_A, \T_B)$, which remains valid for both static entropy density and free streamed energy density initialization.

\begin{figure}
  \begin{tikzpicture}
    \node {\includegraphics{thickness}};
    \node[label=above:{10 fm}] (a) at (0, 1.1) {};
    \node (b) at (0, -1.1) {};
    \draw [<->, semithick] (a) to (b);
  \end{tikzpicture}
  \caption{
    \label{fig:thickness}
    Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
    The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.
  }
\end{figure}

Up to this point we have restricted our attention to a single nucleon-nucleon collision.
Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should, in principle, be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
The \trento\ model readily generalizes from individual nucleon-nucleon collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
Figure~\ref{fig:thickness} shows the effect of adding nucleon substructure to a generic lead nucleus.
Additional fluctuations emerge over the length scale of a nucleon, but the macroscopic geometry of the nucleus is largely unchanged.

@@

\begin{figure}
  \includegraphics{coupling}
  \caption{
    \label{fig:coupling}
    Cartoon of the free streaming approximation for hydrodynamic initialization. The initial state is free streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling).
    This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.
  }
\end{figure}

@@

The parametric energy deposition ansatz in Eq.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
\begin{equation}
  \label{freestream}
  p^\mu \partial_\mu f(x, p) = 0
\end{equation}
through its dependence on the underlying distribution function $f(x, p)$.
Equation~\eqref{freestream}, however, simplifies for a boost-invariant gas of massless noninteracting partons with locally isotropic transverse momentum distributions.
Subject to these assumptions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ equals \cite{Broniowski:2008qk, Liu:2015nwa}
\begin{multline}
  \label{energy-momentum}
  T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau) =\\
  \frac{\tau_0}{\tau} \int_{-\pi}^{\pi} d\phi_p\, \hat{p}^\mu \hat{p}^\nu e(\xv_\perp - (\tau - \tau_0)\hat{\mathbf{p}}_\perp, \eta_s=0, \tau_0),
\end{multline}
where $\hat{p}^\mu = (1, \cos \phi_q, \sin \phi_q, 0)$ and $\hat{\mathbf{p}}_\perp = (\cos \phi_q, \sin \phi_q)$ are momentum unit vectors.
Here we assume that the time $\tau_0 \ll 1~\fmc$ is small and define $\taufs \equiv \tau - \tau_0 \approx \tau$.
We also combine the unknown constant in Eq.~\eqref{eq:gmean} with the pre-factor $\tau_0$ in Eq.~\eqref{energy-momentum} to produce a single normalization factor $\text{Norm} = \text{const} \times \tau_0$ with units of energy.
The solution \eqref{energy-momentum} is then decomposed in hydrodynamic form
\begin{equation}
  \label{hydro-eqn}
  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu},
\end{equation}
where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} = g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue
\begin{equation}
  T^{\mu\nu} u_\nu = e u^\mu.
\end{equation}
The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eq.~\eqref{hydro-eqn}:
\begin{align}
  \Pi &= -\frac{1}{3} \mathrm{Tr}(\Delta_{\mu\nu} T^{\mu\nu}) - P,\\
  \pi^{\mu\nu} &= T^{\mu\nu} - e u^\mu u^\nu + (P + \Pi) \Delta^{\mu\nu}.
\end{align}

@@

\begin{figure}[t]
  \includegraphics{viscosity_dof}
  \caption{
    \label{fig:viscosity_dof}
    Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations. Lines are chosen for illustrative purposes only and do not represent all possible variability.
    For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.
  }
\end{figure}

@@

The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\ym$ describe experimental measurements $\ye$.
The problem involves three distinct components:
\begin{enumerate}[itemsep=0pt, leftmargin=2\parindent]
  \item
    $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
  \item
    $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
  \item
    $E$: the evidence provided by the model $\ym$, the experimental data $\ye$, and their corresponding uncertainties.
\end{enumerate}
As a practical matter, we always assume that hypothesis $H_f$ is correct, meaning there are no glaring flaws in our chosen theoretical framework.
This is a significant assumption; the application of hydrodynamic simulations to small collision systems is speculative, and our conclusions are conditional on the framework making sense.
